---
title: "Down-regulation of MALAT1 is a hallmark of tissue and peripheral proliferative T cells in COVID-19"
author: "Shoumit Dey"
date: "23/03/2021"
output:
  word_document: default
  html_document: default
---

```{r message=FALSE, warning=FALSE, include=FALSE, results='hide'}
library(Seurat)
#library(SeuratDisk)
library(dplyr)
library(spatstat)
library(gplots)
library(corrplot)
library(AUCell)
library(SingleCellExperiment)
#library(customSeuratAnalysis)
#library(tidyverse)
exp_path<-"."
setwd(exp_path)
```
Source: Lung transplantation for patients with severe COVID-19
BY ANKIT BHARAT, MELISSA QUERREY, NIKOLAY S. MARKOV, SAMUEL KIM, CHITARU KURIHARA, RAFAEL GARZA-CASTILLON, ADWAIY MANERIKAR, ALI SHILATIFARD, RADE TOMIC, YULIYA POLITANSKA, HIAM ABDALA-VALENCIA, ANJANA V. YELDANDI, JON W. LOMASNEY, ALEXANDER V. MISHARIN, G. R. SCOTT BUDINGER

SCIENCE TRANSLATIONAL MEDICINE 16 DEC 2020

Some patients with severe COVID-19 develop end-stage pulmonary fibrosis for which lung transplantation may be the only treatment.


```{r load & QC}
GSE158127_T<-readRDS(paste0(exp_path,"/all_T_cells_Bharat_only.rds"))

GSE158127_T <- PercentageFeatureSet(GSE158127_T, pattern = "^MT-", col.name = "percent.mt")

VlnPlot(GSE158127_T, features = c("percent.mt","nCount_RNA","nFeature_RNA"), pt.size = 0.01, group.by = "Cell.Type")
  
FeatureScatter(GSE158127_T, feature1 = "nFeature_RNA", feature2 = "nCount_RNA")

# GSE158127_T <- subset(GSE158127_T, subset = percent.mt < 10 & nFeature_RNA < 2300 & nCount_RNA < 3300 & nCount_RNA > 800)
# 
# VlnPlot(GSE158127_T, features = c("percent.mt","nCount_RNA","nFeature_RNA"), pt.size = 0.01, group.by = "Cell.Type")
#   
# FeatureScatter(GSE158127_T, feature1 = "nFeature_RNA", feature2 = "nCount_RNA")



```


```{r re-analyse}

#Scaling the data - linear transformation
all.genes <- rownames(GSE158127_T)

GSE158127_T <- FindVariableFeatures(GSE158127_T, selection.method = "vst")

GSE158127_T <- ScaleData(GSE158127_T, features = all.genes)
#GSE158127_T[["scale.data"]] <- GSE158127_T[["RNA"]]

GSE158127_T <- RunPCA(GSE158127_T, features = VariableFeatures(object=GSE158127_T))
#
print(GSE158127_T[["pca"]], dims = 1:5, nfeatures = 5)
#
VizDimLoadings(GSE158127_T, dims = 1:2, reduction = "pca")
#
DimPlot(GSE158127_T, reduction = "pca")

ElbowPlot(GSE158127_T)

GSE158127_T <- FindNeighbors(GSE158127_T, dims = 1:5)
#
GSE158127_T <- FindClusters(GSE158127_T, resolution = 0.2)

GSE158127_T <- RunUMAP(GSE158127_T, dims = 1:5)

DimPlot(GSE158127_T, label = TRUE, label.box = TRUE) 
VlnPlot(GSE158127_T, features = c("MALAT1","MKI67"))
############################################################
```

```{r check_T_cell_markers}
GSE158127_T <- fn_add_t_cell_modules(GSE158127_T, "RNA")
FeaturePlot(GSE158127_T, reduction = "umap", features = c("CD8_Tn1","CD8_Tem1","CD8_Trm1","CD8_Temra1","CD8_Tex1","CD8_Tmait1","CD8_Tgd1")) + DimPlot(GSE158127_T, label = TRUE, label.box = TRUE)

FeaturePlot(GSE158127_T, reduction = "umap", features = c("CD4_Tn1","CD4_Tem1","CD4_Th11","CD4_Th171","CD4_Treg1"), combine = TRUE) + DimPlot(GSE158127_T, label = TRUE, label.box = TRUE)
```


```{r AU_cell prep}
#for graph saving reference
study<-"GSE158127_T"
#get count data
exprMatrix <- GSE158127_T[['RNA']]@data

cellsUMAP<-GSE158127_T@reductions[["umap"]]@cell.embeddings
plot(cellsUMAP, pch=16, cex=.3)
#load gmt file
gmtFile <- "gene_lists_auc.gmt"

geneSets <- getGmt(gmtFile)
#Let's check how many of these genes are in the expression matrix:
geneSets <- subsetGeneSets(geneSets, rownames(exprMatrix)) 
cbind(nGenes(geneSets))

#To ease the interpretation of the tutorial, we will also add the gene-set size into its name:
geneSets <- setGeneSetNames(geneSets, newNames=paste(names(geneSets), " (", nGenes(geneSets) ,"g)", sep=""))

# Random
set.seed(321)
extraGeneSets <- c(
  GeneSet(sample(rownames(exprMatrix), 50), setName="Random (50g)"),
  GeneSet(sample(rownames(exprMatrix), 500), setName="Random (500g)"))

countsPerGene <- apply(exprMatrix, 1, function(x) sum(x>0))
# Housekeeping-like
extraGeneSets <- c(extraGeneSets,
                   GeneSet(sample(names(countsPerGene)[which(countsPerGene>quantile(countsPerGene, probs=.95))], 100), setName="HK-like (100g)"))

geneSets <- GeneSetCollection(c(geneSets,extraGeneSets))
names(geneSets)

gc()
```


```{r AU_Cell build rankings, calculate AUC & explore thresholds }
#build rankings
cells_rankings <- AUCell_buildRankings(exprMatrix, nCores=1, plotStats=TRUE)
cells_rankings
#calculate AUC
cells_AUC <- AUCell_calcAUC(geneSets, cells_rankings)
#explore thresholds
par(mfrow=c(3,3)) 
  cells_assignment <- AUCell_exploreThresholds(cells_AUC, plotHist=TRUE, assign=TRUE) 
  warningMsg <- sapply(cells_assignment, function(x) x$aucThr$comment)
  warningMsg[which(warningMsg!="")]
```

```{r select thresholds and graph}
selectedThresholds <- getThresholdSelected(cells_assignment)
selectedThresholds[1] <-  0.39
selectedThresholds[2] <-  0.25
selectedThresholds[3] <-  0.3
selectedThresholds[4] <-  0.32
selectedThresholds[5] <-  0.3
selectedThresholds[6] <-  0.25
selectedThresholds[7] <-  0.02
selectedThresholds[8] <-  0.035
selectedThresholds[9] <-  0.35
# selectedThresholds[10] <-  0.5
# selectedThresholds[11] <-  0.02
# selectedThresholds[12] <-  0.04
# selectedThresholds[13] <-  0.4

type<-"selected"

#pdf("D:/GoogleDrive/spatial_leishmania/spatial/SingleCellAnalysis/human/Covid-19/combined/Fig4B.pdf")
  par(mfrow=c(2,3))
  AUCell_plotTSNE(tSNE=cellsUMAP, exprMat=exprMatrix, cellsAUC=cells_AUC[1:9,], thresholds=selectedThresholds)
#dev.off()

save_auc_values<-data.frame(t(cells_AUC@assays@data@listData[["AUC"]]))
save_auc_values$barcodes<-rownames(save_auc_values)
#save_auc_values$CD4_all_anti_124g<-save_auc_values$CD4_all_anti..124g.

GSE158127_T <- AddMetaData(GSE158127_T, save_auc_values)


```


```{r exploring_proliferative_cells}
#Take cells that have a high AUC score
GSE158127_T_AUC_pos <- subset(GSE158127_T, subset = CD4_CD8_acorr_common..88g. > 0.39)

#Scaling the data - linear transformation
all.genes <- rownames(GSE158127_T_AUC_pos)

GSE158127_T_AUC_pos <- FindVariableFeatures(GSE158127_T_AUC_pos, selection.method = "vst")

GSE158127_T_AUC_pos <- ScaleData(GSE158127_T_AUC_pos, features = all.genes)
#GSE158127_T_AUC_pos[["scale.data"]] <- GSE158127_T_AUC_pos[["RNA"]]

GSE158127_T_AUC_pos <- RunPCA(GSE158127_T_AUC_pos, features = VariableFeatures(object=GSE158127_T_AUC_pos))
#
print(GSE158127_T_AUC_pos[["pca"]], dims = 1:5, nfeatures = 5)
#
VizDimLoadings(GSE158127_T_AUC_pos, dims = 1:2, reduction = "pca")
#
DimPlot(GSE158127_T_AUC_pos, reduction = "pca")

ElbowPlot(GSE158127_T_AUC_pos)

GSE158127_T_AUC_pos <- FindNeighbors(GSE158127_T_AUC_pos, dims = 1:15)
#
GSE158127_T_AUC_pos <- FindClusters(GSE158127_T_AUC_pos, resolution = 0.4)

GSE158127_T_AUC_pos <- RunUMAP(GSE158127_T_AUC_pos, dims = 1:15)

DimPlot(GSE158127_T_AUC_pos, label = TRUE, label.box = TRUE) 
```

```{r doublets_by_simulation}
library(BiocSingular)
library(scDblFinder)
set.seed(100)

GSE158127_T_AUC_pos_sce <- as.SingleCellExperiment(GSE158127_T_AUC_pos)

library(scuttle)
sce <- logNormCounts(GSE158127_T_AUC_pos_sce)

library(scran)
dec <- modelGeneVar(GSE158127_T_AUC_pos_sce)
hvgs <- getTopHVGs(dec, n=1000)

library(scater)
set.seed(1002)
GSE158127_T_AUC_pos_sce <- runPCA(GSE158127_T_AUC_pos_sce, ncomponents=10, subset_row=hvgs)
GSE158127_T_AUC_pos_sce <- runTSNE(GSE158127_T_AUC_pos_sce, dimred="PCA")

# Setting up the parameters for consistency with denoisePCA();
# this can be changed depending on your feature selection scheme.
dbl.dens <- computeDoubletDensity(GSE158127_T_AUC_pos_sce, subset.row=hvgs)
   # d=ncol(reducedDim(GSE158127_T_AUC_pos_sce)))
plotTSNE(GSE158127_T_AUC_pos_sce, colour_by=I(log1p(dbl.dens)))

GSE158127_T_AUC_pos_sce$DoubletScore <- dbl.dens

plotTSNE(GSE158127_T_AUC_pos_sce, colour_by="DoubletScore")

dbl.calls <- doubletThresholding(data.frame(score=dbl.dens),
    method="griffiths", returnType="call")
summary(dbl.calls)

plotColData(GSE158127_T_AUC_pos_sce,  y="DoubletScore", colour_by=I(dbl.calls))

plotColData(GSE158127_T_AUC_pos_sce, y="DoubletScore", colour_by=I(dbl.calls))

extract_meta <- data.frame(GSE158127_T_AUC_pos_sce@colData)
#extract_meta <- data.frame(extract_meta[, "DoubletScore"])

GSE158127_T_AUC_pos <- AddMetaData(GSE158127_T_AUC_pos, metadata = extract_meta)

plotColData(GSE158127_T_AUC_pos_sce, y="DoubletScore", colour_by=I(dbl.calls)) + FeaturePlot(GSE158127_T_AUC_pos, features = "DoubletScore")

GSE158127_T_AUC_pos <- subset(GSE158127_T_AUC_pos, subset = DoubletScore < 2.5)
DimPlot(GSE158127_T_AUC_pos, label = TRUE, label.box = TRUE) 
```

```{r check_T_cell_markers}
GSE158127_T_AUC_pos <- fn_add_t_cell_modules(GSE158127_T_AUC_pos, "RNA")
FeaturePlot(GSE158127_T_AUC_pos, reduction = "umap", features = c("CD8_Tn1","CD8_Tem1","CD8_Trm1","CD8_Temra1","CD8_Tex1","CD8_Tmait1","CD8_Tgd1")) + DimPlot(GSE158127_T_AUC_pos, label = TRUE, label.box = TRUE)

FeaturePlot(GSE158127_T_AUC_pos, reduction = "umap", features = c("CD4_Tn1","CD4_Tem1","CD4_Th11","CD4_Th171","CD4_Treg1"), combine = TRUE) + DimPlot(GSE158127_T_AUC_pos, label = TRUE, label.box = TRUE)
```

```{r markers_by_cluster}

    
    GSE158127_T_AUC_pos <- SetIdent(GSE158127_T_AUC_pos, value = GSE158127_T_AUC_pos$seurat_clusters)
    
    GSE158127_T_AUC_pos.cluster.markers <- FindAllMarkers(GSE158127_T_AUC_pos, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
    

    
    
    write.csv(GSE158127_T_AUC_pos.cluster.markers, paste0(exp_path, "GSE158127_T_AUC_pos_all_markers_per_ident.csv"))
    
    VlnPlot(GSE158127_T_AUC_pos, features = "MALAT1") + NoLegend()
    


```


```{r prep_figure4_figures_here}
#Fig4
hist(GSE158127_T$CD4_CD8_acorr_common..88g., 
     col="lightblue1",
     freq = FALSE,
     main = "Area under curve - gene list",
     xlab="AUC values (CD4/CD8 MALAT1 acorr - 88 genes)", 
     ylab = "Density")
lines(density(GSE158127_T$CD4_CD8_acorr_common..88g.))
polygon(density(GSE158127_T$CD4_CD8_acorr_common..88g.), col=rgb(1,0,1,.2))
        

#Fig4b
auc_plot <- FeaturePlot(GSE158127_T, features = c("CD4_CD8_acorr_common..88g."), min.cutoff = selectedThresholds[1], cols = c("grey", "red")) + ggtitle(label = "Area under curve", subtitle = "CD4/CD8 MALAT1 acorr (88 genes)") + NoLegend()


#Fig4c
FeaturePlot(GSE158127_T, features = c("MKI67","MALAT1"), ncol = 1) 

#Fig4d
# top10 <- GSE158127_T_AUC_pos.cluster.markers %>%
#   group_by(cluster) %>%
#     top_n(n = 10, wt = avg_log2FC) 
    GSE158127_T_AUC_pos.cluster.markers %>%
      group_by(cluster) %>%
        slice_max(n = 10, order_by = avg_log2FC) -> top10


DoHeatmap(GSE158127_T_AUC_pos, features = top10$gene) + NoLegend() + theme(text = element_text(size = 12))

#Fig4c
FeaturePlot(GSE158127_T_AUC_pos, features = c("MKI67","MALAT1"), ncol = 1) 

#Figure 4 supplementary

p <- FeaturePlot(GSE158127_T_AUC_pos, features = c("TRBC2","CD8A","CD4"), combine = FALSE)
for(i in 1:length(p)) {
  p[[i]] <- p[[i]] + NoLegend() #+ NoAxes()
}
next_slot <- length(p)+1
p[[next_slot]] <- DimPlot(GSE158127_T_AUC_pos, label = TRUE, label.box = TRUE) + NoLegend()
cowplot::plot_grid(plotlist = p, ncol = 2) 

#Figure 4
VlnPlot(GSE158127_T_AUC_pos, features = "MALAT1", ncol = 1) +  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(), panel.background=element_rect(fill = "white") )

#CellCycleScoring for entire object
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
GSE158127_T <- CellCycleScoring(GSE158127_T, s.features = s.genes, g2m.features = g2m.genes)

FeatureScatter(GSE158127_T, feature1 = "S.Score", feature2 = "G2M.Score")

all_cells <- FeatureScatter(GSE158127_T, feature1 = "S.Score", feature2 = "MALAT1", group.by = "Phase") + NoLegend() +
  FeatureScatter(GSE158127_T, feature1 = "G2M.Score", feature2 = "MALAT1", group.by = "Phase")

VlnPlot(GSE158127_T, features = "MALAT1", group.by = "Phase" ) + NoLegend()

FeatureScatter(GSE158127_T, feature1 = "MKI67", feature2 = "MALAT1")

#CellCycleScoring for AUC pos object
GSE158127_T_AUC_pos <- CellCycleScoring(GSE158127_T_AUC_pos, s.features = s.genes, g2m.features = g2m.genes)

FeatureScatter(GSE158127_T_AUC_pos, feature1 = "S.Score", feature2 = "G2M.Score")

auc_pos_cells <- FeatureScatter(GSE158127_T_AUC_pos, feature1 = "S.Score", feature2 = "MALAT1", group.by = NULL) + NoLegend() +
  FeatureScatter(GSE158127_T_AUC_pos, feature1 = "G2M.Score", feature2 = "MALAT1", group.by = NULL) +  theme(axis.text.y = element_blank(), axis.title.y = element_blank(), axis.ticks.y = element_blank(), panel.background=element_rect(fill = "white") )

FeatureScatter(GSE158127_T_AUC_pos, feature1 = "MKI67", feature2 = "MALAT1")
DimPlot(GSE158127_T_AUC_pos, group.by = "Sample.Name") + DimPlot(GSE158127_T_AUC_pos, group.by = "COVID.19")
#vlnplot malat
#idents = c("S","G2M")
GSE158127_T <- SetIdent(GSE158127_T, value = GSE158127_T$Phase)
GSE158127_T_AUC_pos <- SetIdent(GSE158127_T_AUC_pos, value = GSE158127_T$Phase) 
parent_obj <- VlnPlot(GSE158127_T, features = c("MALAT1"), ncol = 1) + theme(axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(), panel.background=element_rect(fill = "white") ) 
daughter_obj <- RidgePlot(GSE158127_T_AUC_pos, features = "MALAT1", group.by = "Phase", idents = c("S","G2M")) +  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(), panel.background=element_rect(fill = "white") )
GSE158127_T <- SetIdent(GSE158127_T, value = GSE158127_T$seurat_clusters)
GSE158127_T_AUC_pos <- SetIdent(GSE158127_T_AUC_pos, value = GSE158127_T_AUC_pos$seurat_clusters) 
parent_obj 
daughter_obj

DimPlot(GSE158127_T_AUC_pos, group.by = "Phase") + ggtitle(label = "Cellcycle Phase", subtitle = "Bharat et al - AUC>0.39 only") + theme(plot.subtitle=element_text(hjust=0.5))


```

```{r test_chunk}
require(stats)

gene_table <- data.frame(t(GSE158127_T_AUC_pos@assays[["RNA"]]["MALAT1" ,]))
gene_table$barcode <- rownames(gene_table)
meta_table <- GSE158127_T_AUC_pos@meta.data
meta_table$barcode <- rownames(GSE158127_T_AUC_pos@meta.data)


#JOIN gene and meta table
gene_meta <- sqldf("SELECT gene_table.barcode, MALAT1, meta_table.'G2M.Score' as G2Mscore, meta_table.'S.Score' as Sscore
                    FROM gene_table
                    INNER JOIN meta_table
                    ON gene_table.barcode = meta_table.barcode")

cor.test(gene_meta$MALAT1, gene_meta$G2Mscore)
cor.test(gene_meta$MALAT1, gene_meta$Sscore)

write.csv(gene_meta, paste0(exp_path,"CellCycleScoring_GSE158127_T_AUC_pos.csv"))

# sp <- FeatureScatter(GSE158127_T_AUC_pos, feature1 = "S.Score", feature2 = "MALAT1", group.by = NULL) + NoLegend()
# reg<-lm(Sscore ~ MALAT1, data = gene_meta)
# reg
# coeff=coefficients(reg)
# coeff
# # Equation of the line : 
# eq = paste0("y = ", round(coeff[2],1), "*x + ", round(coeff[1],1))
# # Plot
# sp + geom_abline(intercept = coeff[1], slope = coeff[2], color="red",linetype="dashed", size=1.5)#+
#  ggtitle(eq)
# Change line type, color and size
#sp + geom_abline(intercept = 37, slope = -5, color="red", 
   #              linetype="dashed", size=1.5)#+
  #ggtitle(eq)
```